Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery.

### 2024-03-01 docs/index.qmd
Pierre comments

Coastal areas are vital hotshots for marine biodiversity, with intertidal seagrass meadows playing a crucial role at the interface between land and ocean [@unsworth2022]. These meadows offer a myriad of ecosystem services to humanity, including limitation of coastline erosion, reducing the risk of eutrophication, carbon sequestration, and oxygen production. They serve as vital habitats for a diverse array of marine and terrestrial species, providing living, breeding, and feeding grounds [@gardner2018 ; @Zoffoli2022 ; @jankowska2019]. Due to their proximity to human activities, seagrass meadows are directly exposed to and impacted by anthropogenic pressures. Global regression and fragmentation are currently observed due to diseases, disasters, coastal urbanization, sea reclamation, as well as fishing activities, dredging, sea level rise, coastal erosion, competition with alien species, and reduction in water quality [@nguyen2021 ; @soissons2018 ; @orth2006 ; @lin2018 ; @duffy2019]. While improvements in water quality have been recently reported in European sites, allowing an overall recovery of seagrass ecosystems at the local scale, many other coastal waters worldwide are still subjected to strong eutrophication processes [@deSantos2019 ; @Zoffoli2021]. Coastal eutrophication has been associated to anomalous accumulation of green macroalgae, the so-called green tides. Green tides produce shade and suffication over seagrass individuals, thus threatening the health of seagrass ecosystems [@Duarte2002 ; @wang2022].

The importance of seagrass meadows and the variety of ecosystem services they provide have led to the enhancement of global and regional monitoring programs for systematically surveying different Essential Oceanic Variable [@Miloslavich2018] as seagrass coverage and composition; as well as Essential Biodiversity Variable [@Pereira2013] such as seagrass taxonomic diversity; species distribution, population abundance, and phenology. Monitoring programs also prioritize the identification of threats to these ecosystems, particularly during early stages, to facilitate effective mitigation actions. Traditionally, these ecological parameters have been quantified through in situ measurements, although this approach faces several constraints over intertidal zones. Intertidal meadows are only partially exposed during low tide and can be situated in difficult-to-reach mudflats, potentially leading to inaccurate and limited estimations with conventional sampling techniques [@nijland2019]. However, satellite data have been proven effective in complenmenting in situ surveys, allowing for the rapid and consitent retrieval of EOV's over extensive seagrass meadows. [@Zoffoli2021 ; @xu2021 ; @Traganos2018 ; @coffer2023]

Satellite remote sensing offers the advantage of acquiring large-scale data in real-time but presents its inherent challenges. Free access satellite data (e.g., Sentinel-2 and Landsat8/9) provide relatively low spatial resolution data (10 - 30 m) across a limited number of spectral bands. These characteristics can be a limitation to accurately discriminating seagrass from others co-existing macrophytes over the meadow. Chlorophyceae (Green Algae) and marine Magnoliopsida (Seagrass) share the same pigment composition [@ralph2002 ; @Douay2022]. As a result, their respective spectral signatures can be considered similar by a non-expert observer [@Davies2023 ; @bannari2022]. Recently, using a hyperspectral library, @Davies2023 demonstrated that the spectral resolution of Sentinel-2, might be enough for the discrimination between Magnoliopsida and Chlorophyceae. However, green tide events occur at small spatial scales that are not observable using satellite imagery [@tuya2013], especially during the initial stage of the event.

Remote sensing drone acquisitions are presented as a tool that can potentially fill gaps left by satellite and in situ data. Drone can cover large expanses while recording imagery at significantly higher spatial resolutions than satellite (pixel size from cm to mm) and still capturing data at multi-spectral resolution [@fairley2022drone ; @oh2017use]. The versatility of drones allows for their application across a diverse thematic range , from coastal zone management [@adade2021 ; @casella2020 ; @angnuureng2022] to mapping the spatial distribution of species [@joyce2023 ; @tallam2023 ; @Roca2022 ; @Roman2021 ; @Brunier2022Topographic]. However, when applied to coastal habitat mapping, many studies showcase their findings with study case limited to a single flight, restricting the generalizability of their application to other sites [@Roman2021 ; @collin2019improving ; @rossiter2020uav ; @Brunier2022Topographic]. This study aimed to analyze the potential of a drone equipped with a multispectral sensor for mapping intertidal macrophytes, with a particular focus on discriminating Magnoliopsida and Chlorophyceae. Ten drone flights were performed over soft-bottom intertidal areas along two European countries (France and Portugal), covering a wide range of habitats, from monospecific seagrass meadows to meadows mixed with green or red algae. A deep learning algorithm was trained and validated for macrophyte discrimination, emphasizing applicability across diverse sites without a loss of accuracy in predictions.

@@

```{r Figure1}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: >
#|   Sites where the different drone flights were made. A: Gulf of Morbihan, B: Bourngeuf Bay, C: Ria de Aveiro Lagoon. Light green represent terrestrial areas whereas darkgreen represent intertidal areas. 
#| label: fig-map
#| fig-width: 7
#| fig-height: 7
#| out-width: "100%"

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a Micasense RedEdge Dual MX multispectral camera was flown to take 1.2 million pixel reflectance photographs in 10 spectral bands, from blue (444 nm) to near infrared (840 nm). An angle of 90° has been maintained between the sun and the drone's route to guarantee uniform lighting conditions between flight lines. A side overlap of 70% and a front overlap of 80% between each image has been set. Images have been calibrated in reflectance using a calibration panel reflective at \~50% provided by the camera's manufaturer and a Downwelling Light Sensor (DLS2) that was utilized to acquire irradiance data during the flight. A Structure-from-Motion photogrammetry software (Agisoft Metashape) has been used to process images to obtain multispectral orthomosaics of each flight. The workflow for orthomosaicking has remained consistent for every flight. First, tying key points were detected inside of each image and between overlapping images in order to obtain a sparse point cloud. This cloud was cleaned using reprojection accuracy metric in order to remove noisy points. A dense point cloud has been then produced using a structure from motion algorithm. A surface interpolation of this dense point cloud has been made to obtain a Digital Surface Model (DSM), used to reconstruct the multispectral ortho-image. Across all the sites, flights were made at two different altitudes : 12 m or/and 120 m. Low altitude drone flights produce ortho-images with a very high spatial resolution (8 mm per pixel), making it simple to visually distinguish between the various types of vegetation. High altitude fights on the other hand allow to cover large areas and produced images with a pixel size of 80 mm (Table 1).

```{r Table1}
#| echo: false
#| warning: false
#| tbl-cap: List of drone Flight, summarising the date, the altitude and the purpose of each flight.
#| cache: true

@@

Before each flight, targets used as Ground Control Points (GCPs) were dispersed and georeferenced with a Trimble © Geo XH 6000 differential GPS (dGPS). GCPs were used to correct georeferencing imprecision of orthomosaics with an horizontal and vertical accuracy of 10cm. dGPS was also used to georeference quardrats of 0.25cm² used to assess the presence or absence of 5 key classes of vegetation : Bacillariophyceae (Diatoms forming large biofilm over the sediment during low tide), Phaeophyceae (Brown algae), Magnoliopsida (Seagrass), Chlorophyceae (Green algae) and Rhodophyceae (Red algae) (@fig-vegetation). Picture of each quadrat were taken and uploaded on the online on the Global Biodiversity Information Facility (GBIF) platform [@BedeGbif]. Each photograph was also processed to estimate the percent cover of each type of vegetation. Georeferenced hyperspectral reflectance signatures of each vegetation class were also sampled using an ASD FieldSpec HandHeld 2 spectroradiomete, which acquires reflectance between 325 and 1075 nm, with 1 nm of spectral resolution. The methodological scheme followed in this study is presented in @fig-workflow.

@@

```{r Figure2 Plot}
#| cache: true
#| echo: false
#| warning: false
#| fig-cap: Class of vegetation used to train the CNN model and their standardised spectral signature
#| label: fig-vegetation
#| fig-width: 10
#| fig-height: 7
#| out-width: "100%"

@@

```{r Figure3}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: "Schematic representation of the workflow. Diamonds represent input or output data, and rectangles represent Python processing algorithms. The overall workflow of this study is divided into two distinct parts based on the spatial resolution of the drone flights: high-resolution flights were utilized for training and prediction of the CNN model, whereas low-resolution flights were solely employed for prediction purposes."
#| label: fig-workflow
#| fig-width: 10
#| fig-height: 7
#| out-width: "100%"

@@

The spectral similarities of reflectance signatures between intertidal green macrophytes (Magnoliopsida and Chlorophyceae) make their discrimination challenging using simple classification algorithms (@fig-vegetation F). To overcome this challenge, a deep learning classification method was developed, trained, validated, and applied to each drone flight (@fig-workflow).

@@

A dataset containing photo-interpreted drone reflectance pixels was built to train a Convolutional Neural Network (CNN) model with 2 hidden layers. The training pixels were categorized into 7 different classes, representing the various habitats encountered at the different study sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and Rhodophyceae. Only low-altitude flights (Table 1) were used for training purposes because of their 8 mm spatial resolution allowing to avoid spectral sub-pixel mixing and to accurately visually differentiate various vegetation classes. More than `r my_comma(round(sum(df_training$Training_Pixels),-3))` pixels at 8 mm resolution from the 3 training flights have been used to train the model (Table 2). Twenty one variables were used by the model as predictors: 10 raw spectral bands of the Micasense RedEdge Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same 10 spectral bands standardized using a min/max transformation ([@eq-std] ; @Cao2017) and the NDVI (@eq-ndvi). Standardisation of spectral bands is used to eliminate the scaling differences between spectra and to limit the effect of biomass on the shape of the spectra [@Douay2022 ; @Davies2023].

@@

The CNN model was applied to all the 9 flights at both 12 and 120 m of altitude. In situ information on georeferenced class type and percent cover collected during each flight has been used to assess the CNN accuracy. These images were used to construct a validation dataset indicating the presence or absence of each class. Additionally to the quadrat-based validation dataset, polygons of each class were photo interpreted in order to increase the number of pixels of the validation dataset. A confusion matrix, along with precision metrics such as the global accuracy, the sensitivity, the specificity, and the Kappa coefficient, was generated for each of the study sites. Finally each validation matrix has been merge to create a unique matrix of the model. A total of `r my_comma(round(count_validation_pixel,-3))` pixels has been used to globally validate the CNN model.

@@

A total of 9 prediction maps corresponding to the 9 drone flights were obtained. Each prediction map is associated with a probability map, indicating the probability of the selected class for every pixel. The low-altitude flight conducted in Gafanaha, Portugal, represents the site with the highest complexity (@fig-GafLow). Among the 5 vegetation classes on which the CNN model was trained, 4 were present on this site. On this site, there is a mixture of Chlorophyceae and Rhodophyceae over the seagrass meadow. This is also where Bacillariophyceae is most abundant. Although the seagrass bed is solely composed of *Zostera noltei*, various colors can be observed: dark green (indicating healthy beds with 100% coverage) and whitish/brown (indicating beds where leaves are dying). Regardless of the color, the meadow is predicted as Magnoliopsida by the CNN model.

@@

The high-altitude flight over Gafanha covered nearly 1 km² in total (@fig-GafHigh). A channel delineating a small island was present in the flight, but it has been masked in the prediction map due to glinting and misclassification caused by the turbid shallow water. Most of the intertidal area has been classified as Magnoliopsida by the model, even though there is discoloration of seagrass blades. Only a few pixels have been classified as Chlorophyceae at this scale. Furthermore, the area that was classified as Bacillariophyceae in the low-altitude flight remains mostly classified as such in the high-altitude flight, but some pixels of it are now classified as Magnoliopsida. The patches of Rhodophyceae are still well classified. *Spartina sp.* was present in the northern part of the flight, near the land, and has been classified either as Magnoliopsida or as Phaeophyceae.

@@

The High altitude flight made in the inner lagoon of the Ria de Aveiro is the largest of all flights, covering almost 1.5 km² (@fig-Boat). On this site, only seagrass and red algae where seen on the field. The classification follows the same pattern, with a patchy Magnoliopsida meadow mixed on the eastern part of the flight with Rhodophyceae. As shown in the zoom (@fig-Boat), the edges of the meadow can be colonised by Chlorophyceae.

@@

The flight over L'Epine in Noirmoutier Island, France (@fig-Dike) was conducted near a dike crossing the northern part of the flight from west to east. Alongside this dike, brown algae attached to rocks and stranded green algae could be found. Despite the high mixture between Chlorophyceae and Magnoliopsida in this flight, these two classes are well discriminated by the classifier.

@@

A total amount of `r my_comma(round(count_validation_pixel,-3))` pixels have been used to validate the Neural Network classifier. The site with the least amount of validation data is `r df_summarize[which(df_summarize$n == min(df_summarize$n)),]$Site` with a total of `r min(df_summarize$n)` pixels whereas `r df_summarize[which(df_summarize$n == max(df_summarize$n)),]$Site` is the site with the most amount of validation data with `r max(df_summarize$n)` pixels. The global accuracy of the model is `r round(validation_metrics$overall[[1]]*100,2)`% and a Kappa coefficient of `r round(validation_metrics$overall[[2]],2)` (@fig-Validation). The least performing site is `r BySite_Metrics %>% filter(metric == "Accuracy") %>% filter(value == min(value)) %>% pull(Site)` with an accuracy of `r round(BySite_Metrics %>% filter(metric == "Accuracy") %>% filter(value == min(value)) %>% pull(value)*100,2)`% whereas `r BySite_Metrics %>% filter(metric == "Accuracy") %>% filter(value == max(value)) %>% pull(Site)` is the site with most accurate prediction, with a global accuracy of `r round(BySite_Metrics %>% filter(metric == "Accuracy") %>% filter(value == max(value)) %>% pull(value)*100,2)`%. Overall Phaeophyceae, Magnoliopsida, Sediment and Rhodophyceae are well classified with a balanced accuracy of `r round(Overall_metrics_ByClass %>% filter(metric == "Balanced Accuracy") %>% filter(Class == "Phaeophyceae") %>% pull(value),2)`, `r round(Overall_metrics_ByClass %>% filter(metric == "Balanced Accuracy") %>% filter(Class == "Magnoliopsida") %>% pull(value),2)`, `r round(Overall_metrics_ByClass %>% filter(metric == "Balanced Accuracy") %>% filter(Class == "Sediment") %>% pull(value),2)` and `r round(Overall_metrics_ByClass %>% filter(metric == "Balanced Accuracy") %>% filter(Class == "Rhodophyceae") %>% pull(value),2)` respectively. `r Overall_metrics_ByClass %>% filter(metric == "Balanced Accuracy") %>% filter(value == min(value)) %>% pull(Class)` is the least performing class with an accuracy of `r round(Overall_metrics_ByClass %>% filter(metric == "Balanced Accuracy") %>% filter(value == min(value)) %>% pull(value),2)` mainly due to a confusion with Magnoliopsida and Sediment.

@@

The computation of the importance variable of each class indicate which wavelength is mandatory to accurately predict the class (@fig-VIP). Bands at 444 nm, 717 nm and 842 nm of the Micasense camera are important for none of the vegetation classes. The band at 531 nm is the only important predictor for the classifier to accuratly predict chlorophyceae. In fact, at this wavelength, the Chlorophyceae spectra has the maximum reflectance value of all the classes. 531 nm and 740 nm are the most important predictors to predict Phaeophyceae, corresponding to the minimum reflectance value among all the classes. Bands at 475 and 560 nm are the most important predictors for Bacillariophyceae and Rhodophyceae, respectively. 4 predictors, ranging from the Green (560 nm) to the RedEdge (705 nm) are important to predict accurately magnoliopsida.

@@

@fig-upscaling is showing the percent cover of a class required to classify correctly the high altitude flight. When the percent cover of the class is 100 %, big pixels are well classified for all the classes excepted for Bare Sediment, where it's well classified 80% of the time. A vegetative percent cover of at least 80% is need to have all the big pixels well classified, at the exception of Magnoliopsida that needs an higher percent cover (\>90 %) to be well classified. Concerning the probability of each class, a really high Percent cover is needed to confidently predict Bacillariophyceae. To predict Chlorophyceae with a confidence of 0.85, a percent cover of `r round((lm_green$coefficients[[2]]*85+lm_green$coefficients[[1]])*100,0)` % is needed, `r round((lm_magno$coefficients[[2]]*85+lm_magno$coefficients[[1]])*100,0)` % for magnoliopsida, `r round((lm_Rodo$coefficients[[2]]*85+lm_Rodo$coefficients[[1]])*100,0)` % for Rhodophyceae and `r round((lm_MPB$coefficients[[2]]*85+lm_MPB$coefficients[[1]])*100,0)` % for Bacillariophyceae.

@@

::::: columns
::: column
```{r Global Accuracy assesment Plot}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: Sample of @fig-Validation focusing on green macrophytes. The labels inside the matrice indicate the number of pixels.
#| label: fig-ValidationGreen
#| fig-width: 10
#| fig-height: 10
#| out-width: "100%"

@@

::: column
The Micasense RedEdge-MX DUAL camera used in this study meets these two criteria, enabling the classifier to achieve `r round(validation_metrics_green$overall[[1]],2)*100`% of accuracy between these two classes (@fig-ValidationGreen). Even if the pigment compostion of green macrophytes is similar, differences in the spectral shape can still be observed (@fig-vegetation). Several factors can explain these differences such has different pigment concentration en proportion, different cellular organisation and 3D disposition of the plant a whole on the intertidal mudflat [@beach1997vivo, @kirk1994light, @hedley2018influence]. 
:::
:::::

The VIP analysis of the Neural Network model (see Fig. VIP) shows that the 531 nm band is the most important spectral band for accurately identifying Chlorophyceae. In fact, at this wavelength, Chlorophyceae exhibits the highest reflectance among all other classes, highlighting the difference in carotenoid proportion between seagrasses and green algae.

>>>>

Coastal areas are vital hotshots for marine biodiversity, with intertidal seagrass meadows playing a crucial role at the interface between land and ocean [@unsworth2022]. Seagrass meadows provide a myriad of ecosystem services to humanity, including carbon sequestration, oxygen production,protection against sea-level rise and coastline erosion, and limitation of eutrophication. They serve as vital habitats for a diverse array of marine and terrestrial species, providing living, breeding, and feeding grounds [@gardner2018 ; @Zoffoli2022 ; @jankowska2019]. Due to the concentration of human activities in coastal zones, seagrass meadows are directly exposed to and impacted by anthropogenic pressures. Global regression and fragmentation are currently observed due to climate change, diseases, urbanization, land reclamation, dredging, competition with alien species, and reduction in water quality [@nguyen2021 ; @soissons2018 ; @orth2006 ; @lin2018 ; @duffy2019]. While improvements in water quality have been recently reported in Europe, allowing an overall recovery of seagrass ecosystems at local scale, many coastal waters worldwide are still subjected to strong eutrophication processes [@deSantos2019 ; @Zoffoli2021]. Coastal eutrophication has been associated to anomalous accumulation of green macroalgae, the so-called green tides (@devlin2023nutrients ; @devlin2023nutrients). Green tides produce shade and suffication over seagrass individuals, thus threatening the health of seagrass ecosystems [@wang2022].

The importance of seagrass meadows and the variety of ecosystem services they provide have led to the enhancement of both global and regional programs monitoring Essential Oceanic Variable (EOVs) such as seagrass composition [@Miloslavich2018], as well as Essential Biodiversity Variable (EBVs) such as seagrass taxonomic diversity, species distribution, populuation abundance, and phenology [@Pereira2013]. Seagrass monitoring is all the more important as it makes possible to identify which pressure should be addressed in priority to facilitate effective mitigation actions. Traditionally, indicators of seagrass status have been quantified using in situ measurements. The acquisition of field data in intertidal zones is however notoriously challenging. Intertidal seagrass meadows are only partially exposed during low tide and can be situated in difficult-to-reach mudflats, potentially leading to inaccurate and limited estimations with conventional sampling techniques [@nijland2019]. Satellite observations have been proven effective in complementing in situ sampling, allowing for the near real-time and consistent retrieval of seagrass EOVs and EBVs over extensive meadows. [@Zoffoli2021 ; @xu2021 ; @Traganos2018 ; @coffer2023]

While satellite remote sensing provides temporally consistent observations over large spatial scales, it is limited by several constraints. Satellite missions with a high temporal resolution (e.g. daily MODIS observation) are limited by a coarse spatial resolution (\>100m). Mission with a high spatial resolution such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low spectral resolution. The limited number of spectral bands is indeed a challenge to accurately discriminate seagrass from others co-existing macrophytes. In particular Chlorophyceae (green algae) and marine Magnoliopsida (seagrass) share the same pigment composition [@ralph2002 ; @Douay2022]. As a result, their respective spectral signatures can be considered similar by a non-expert observer [@Davies2023 ; @bannari2022]. Recently, using advanced machine-learning algorithms trained with a large hyperspectral library of more then 300 field reflectance measurements, @Davies2023 demonstrated that it was possible to discriminate Magnoliopsida from Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral resolution. The application of this approach to satellite remote sensing remains however to be validated. Moreover patches of green algae can develop at small spatial scales that are not observable using non-commercial satellite imagery [@tuya2013], especially during the initial stage of the event.

Drone can potentially fill the gaps left by satellite remote sensing and in situ measurments, due to their ability to provide spatially-explicit observations at very high spatial resolutions (pixel size from cm to mm) and still capturing data at multi-spectral resolution [@fairley2022drone ; @oh2017use]. The versatility of drones allows for their application across a diverse thematic range , from coastal zone management [@adade2021 ; @casella2020 ; @angnuureng2022] to mapping species distribution [@joyce2023 ; @tallam2023 ; @Roca2022 ; @Roman2021 ; @Brunier2022Topographic]. However, when applied to coastal habitat mapping, most sutdy cases were limited to a single flight, restricting the generalizability of their application over wider geographical scales [@Roman2021 ; @collin2019improving ; @rossiter2020uav ; @Brunier2022Topographic]. The present study aimed at analyzing the potential of multispectral drone remote sensing to map intertidal macrophytes, with a particular focus on discriminating Magnoliopsida and Chlorophyceae. Ten drone flights were performed over soft-bottom intertidal areas along two European countries (France and Portugal), covering a wide range of habitats, from monospecific seagrass meadows to meadows mixed with green or red algae. A deep learning algorithm was trained and validated for macrophyte discrimination, emphasizing applicability across diverse sites without a loss of prediction accuracy.

@@

```{r Figure1}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: >
#|   Sites observed by drone. A: Gulf of Morbihan, B: Bourngeuf Bay, C: Ria de Aveiro Lagoon. Light green represent terrestrial areas whereas darkgreen represent intertidal areas. 
#| label: fig-map
#| fig-width: 7
#| fig-height: 7
#| out-width: "100%"

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a Micasense RedEdge Dual MX multispectral camera was flown to take 1.2 million pixel reflectance photographs in 10 spectral bands, from 444 nm (blue) to 840 nm (near infrared, NIR). An angle of 90° was maintained between the sun and the drone's route to guarantee uniform lighting conditions between flight lines. A side overlap of 70% and a front overlap of 80% between each imagewas set for each flight. A downwelling light sensor (DLS2) was used to acquire irradiance data concomitantly with the camera measurements. Raw data were calibrated in reflectance using a calibration panel reflective at \~50% provided by the manufacturer. A structure-from-motion photogrammetry software (Agisoft Metashape) was used to process images to obtain multispectral orthomosaics of each flight. The workflow for orthomosaicking was the same for every flight. First, tying key points were detected inside of each image and between overlapping images in order to obtain a sparse point cloud. This cloud was cleaned using reprojection accuracy metric in order to remove noisy points. A dense point cloud was then produced using a structure from motion algorithm. A surface interpolation of this dense point cloud was made to obtain a digital surface model (DSM), used to reconstruct the multispectral ortho-image. Across all sites, flights were made at two different altitudes : 12 m or/and 120 m. Low altitude drone flights produce ortho-images with a very high spatial resolution (8 mm per pixel), making it simple to visually distinguish between the various types of vegetation. High altitude fights on the other hand allow to cover large areas and produced images with a pixel size of 80 mm (Table 1).

```{r Table1}
#| echo: false
#| warning: false
#| tbl-cap: List of drone flight, summarising the date, the altitude and the purpose of each flight.
#| cache: true

@@

Before each flight, targets used as Ground Control Points (GCPs) were distributed over the study site and georeferenced with a Trimble © Geo XH 6000 differential GPS (dGPS). GCPs were used to correct georeferencing imprecision of orthomosaics with an horizontal and vertical accuracy of 10cm. dGPS was also used to georeference quadrats of 0.25cm² used to assess the presence or absence of 5 key classes of vegetation : Bacillariophyceae (Benthic diatoms forming large biofilm over the sediment during low tide), Phaeophyceae (brown algae), Magnoliopsida (seagrass), Chlorophyceae (green algae) and Rhodophyceae (red algae) (@fig-vegetation). Pictures of each quadrat were uploaded on the online on the Global Biodiversity Information Facility (GBIF) platform [@BedeGbif], a public open portal to store and share biodiversity data. Each photograph was also processed to estimate the percent cover of each type of vegetation using an image processing software (imageJ). For all quadrat, the hyperspectral reflectance signatures of each vegetation class was measured using an ASD FieldSpec HandHeld 2 spectroradiometer, which acquires reflectance between 325 and 1075 nm, with 1 nm of spectral resolution. The methodological scheme followed in this study is presented in @fig-workflow.

@@

```{r Figure2 Plot}
#| cache: true
#| echo: false
#| warning: false
#| fig-cap: Class of vegetation used to train the Neural Network model and their standardised spectral signature
#| label: fig-vegetation
#| fig-width: 10
#| fig-height: 7
#| out-width: "100%"

@@

```{r Figure3}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: "Schematic representation of the workflow. Diamonds represent input or output data, and rectangles represent Python processing algorithms. The overall workflow of this study is divided into two distinct parts based on the spatial resolution of the drone flights: high-resolution flights were utilized for training and prediction of the Neural Network model, whereas low-resolution flights were solely employed for prediction purposes."
#| label: fig-workflow
#| fig-width: 10
#| fig-height: 7
#| out-width: "100%"

@@

The spectral similarities of the reflectance signatures between intertidal green macrophytes (Magnoliopsida and Chlorophyceae) make their discrimination challenging using simple classification algorithms (@fig-vegetation F). To overcome this challenge, a deep learning classification method was developed, trained, validated, and applied to each drone flight (@fig-workflow).

@@

A dataset containing photo-interpreted drone reflectance pixels was built to train a Neural Network model with 2 hidden layers. The training pixels were categorized into 7 different classes, representing the various habitats encountered at the different study sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and Rhodophyceae. Only low-altitude flights (Table 1) were used for training purposes because their 8 mm spatial resolution allowied to avoid spectral sub-pixel mixing and to accurately differentiate various vegetation classes. More than `r my_comma(round(sum(df_training$Training_Pixels),-3))` pixels at 8 mm resolution from the 3 training flights were used to train the model (Table 2). Twenty one variables were used by the model as predictors: 10 raw spectral bands of the Micasense RedEdge Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same 10 spectral bands standardized using a min/max transformation ([@eq-std] ; @Cao2017) and the Normalized difference vegetation index (NDVI, @eq-ndvi). Standardisation of spectral bands is used to eliminate the scaling differences between spectra and to limit the effect of biomass on the shape of the spectra [@Douay2022 ; @Davies2023].

@@

The model was applied to all flights at both 12 and 120 m of altitude. In situ information on georeferenced class type and percent cover collected during each flight was used to assess the model accuracy. These images were used to construct a validation dataset indicating the presence or absence of each class. Additionally to the quadrat-based validation dataset, polygons of each class were photo interpreted in order to increase the number of pixels of the validation dataset. A confusion matrix, along with precision metrics such as global accuracy, sensitivity, specificity, and Kappa coefficient, was generated for each sites. All validation matrices were then merged to create a unique matrix . Altogether, a total of `r my_comma(round(count_validation_pixel,-3))` pixels was used to validate the model, thus providing a geographically robust validation dataset.

@@

A total of 9 prediction maps corresponding to the 9 drone flights were obtained. Each prediction map is associated with a probability map, indicating the probability of the selected class for every pixel. The low-altitude flight conducted in Gafanaha, Portugal, represents the site with the highest complexity (@fig-GafLow). Among the 5 vegetation classes on which the model was trained, 4 were present on this site. On this site, there is a mixture of Chlorophyceae and Rhodophyceae over the seagrass meadow. This is also where Bacillariophyceae is most abundant. Although the seagrass bed is solely composed of *Zostera noltei*, various colors can be observed: dark green (indicating healthy beds with 100% coverage) and whitish/brown (indicating beds where leaves are dying). Regardless of the color, the meadow is predicted as Magnoliopsida by the model.

@@

The high-altitude flight over Gafanha covered nearly 1 km² in total (@fig-GafHigh). A channel delineating a small island was masked in the prediction map due to sunglint and misclassification caused by the turbid shallow water. Most of the intertidal area has been classified as Magnoliopsida by the model, even though there is discoloration of seagrass blades. Only a few pixels have been classified as Chlorophyceae at this scale. Furthermore, the area that was classified as Bacillariophyceae in the low-altitude flight remains mostly classified as such in the high-altitude flight, though some pixels were classified as Magnoliopsida. Patches of Rhodophyceae were correctly classified. In the nothern part of the scene, near the land limit, patches of *Spartina sp.* were misclassified, either as Magnoliopsida or as Phaeophyceae.

@@

The high altitude flight acquired over the inner lagoon of the Ria de Aveiro is the largest of all flights, covering almost 1.5 km² (@fig-Boat). On this site, only seagrass and red algae where seen on the field. The classification provided consistent results, with a patchy Magnoliopsida meadow mixed with Rhodophyceae on the eastern part of the scene. As shown in the zoom (@fig-Boat), the edges of the meadow can be colonised by Chlorophyceae.

@@

The flight over L'Epine in Noirmoutier Island, France (@fig-Dike) was conducted near a dike crossing the northern part of the scene from west to east. Alongside this dike, brown algae attached to rocks and stranded green algae could be found. Despite the high mixture between Chlorophyceae and Magnoliopsida these two classes were correctly discriminated by the classifier.

@@

A total of `r my_comma(round(count_validation_pixel,-3))` pixels was used to validate the Neural Network classifier. The sites with the lowest and highest number of validation data were `r df_summarize[which(df_summarize$n == min(df_summarize$n)),]$Site` (`r min(df_summarize$n)` pixels) and `r df_summarize[which(df_summarize$n == max(df_summarize$n)),]$Site` (`r max(df_summarize$n)` pixels), respectively. Model global accuracy was `r round(validation_metrics$overall[[1]]*100,2)`% with a Kappa coefficient of `r round(validation_metrics$overall[[2]],2)` (@fig-Validation). The least performing site was `r BySite_Metrics %>% filter(metric == "Accuracy") %>% filter(value == min(value)) %>% pull(Site)` (global accuracy of `r round(BySite_Metrics %>% filter(metric == "Accuracy") %>% filter(value == min(value)) %>% pull(value)*100,2)`%) whereas `r BySite_Metrics %>% filter(metric == "Accuracy") %>% filter(value == max(value)) %>% pull(Site)` was the site with the most accurate prediction (global accuracy of `r round(BySite_Metrics %>% filter(metric == "Accuracy") %>% filter(value == max(value)) %>% pull(value)*100,2)`%). Overall, the classes Phaeophyceae, Magnoliopsida, Sediment and Rhodophyceae were correctly classified with a balanced accuracy of `r round(Overall_metrics_ByClass %>% filter(metric == "Balanced Accuracy") %>% filter(Class == "Phaeophyceae") %>% pull(value),2)`, `r round(Overall_metrics_ByClass %>% filter(metric == "Balanced Accuracy") %>% filter(Class == "Magnoliopsida") %>% pull(value),2)`, `r round(Overall_metrics_ByClass %>% filter(metric == "Balanced Accuracy") %>% filter(Class == "Sediment") %>% pull(value),2)` and `r round(Overall_metrics_ByClass %>% filter(metric == "Balanced Accuracy") %>% filter(Class == "Rhodophyceae") %>% pull(value),2)` respectively. `r Overall_metrics_ByClass %>% filter(metric == "Balanced Accuracy") %>% filter(value == min(value)) %>% pull(Class)` was the least performing class (accuracy of `r round(Overall_metrics_ByClass %>% filter(metric == "Balanced Accuracy") %>% filter(value == min(value)) %>% pull(value),2)` ) mainly due to the confusion between Magnoliopsida and Sediment.

@@

The computation of the variable importance makes it possible to identify which wavelengths are the most useful for class prediction (@fig-VIP). The spectral bands at 444 , 717 and 842 nm of the Micasense camera are important for none of the vegetation classes. The band at 531 nm is the only important predictor for the classifier to accurately predict chlorophyceae. In fact, at this wavelength, the Chlorophyceae spectra has the maximum reflectance among of all vegetation classes. The bands at 531 and 740 nm are the most important predictors for Phaeophyceae, corresponding to the minimum reflectance among all classes. Bands at 475 and 560 nm are the most important predictors for Bacillariophyceae and Rhodophyceae, respectively. Four predictors, ranging from the green (560 nm) to the RedEdge (705 nm) bands are important to accurately predict magnoliopsida.

@@

Using the very high precision low altitude flight, we determined the minimal percent cover required to classify correctly classify a given class using the high altitude flight (@fig-upscaling). When the percent cover of the class is 100 %, big pixels are well classified for all the classes excepted for Bare Sediment, where it's well classified 80% of the time. A vegetative percent cover of at least 80% is need to have all the big pixels well classified, at the exception of Magnoliopsida that needs an higher percent cover (\>90 %) to be well classified. Concerning the probability of each class, a really high Percent cover is needed to confidently predict Bacillariophyceae. To predict Chlorophyceae with a confidence of 0.85, a percent cover of `r round((lm_green$coefficients[[2]]*85+lm_green$coefficients[[1]])*100,0)` % is needed, `r round((lm_magno$coefficients[[2]]*85+lm_magno$coefficients[[1]])*100,0)` % for magnoliopsida, `r round((lm_Rodo$coefficients[[2]]*85+lm_Rodo$coefficients[[1]])*100,0)` % for Rhodophyceae and `r round((lm_MPB$coefficients[[2]]*85+lm_MPB$coefficients[[1]])*100,0)` % for Bacillariophyceae.

@@

::: columns
::: column
```{r Global Accuracy assesment Plot}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: Sample of @fig-Validation focusing on green macrophytes. The labels inside the matrice indicate the number of pixels.
#| label: fig-ValidationGreen
#| fig-width: 10
#| fig-height: 10
#| out-width: "100%"

@@

::: column
The Micasense RedEdge-MX DUAL camera used in this study meets these two criteria, enabling the classifier to achieve `r round(validation_metrics_green$overall[[1]],2)*100`% of accuracy between these two classes (@fig-ValidationGreen). Even if the pigment compostion of green macrophytes is similar, differences in the spectral shape can still be observed (@fig-vegetation). Several factors can explain these differences such has different pigment concentration en proportion, different cellular organisation and 3D disposition of the plant a whole on the intertidal mudflat [@beach1997vivo, @kirk1994light, @hedley2018influence].
:::
:::

The VIP analysis of the Neural Network model (see Fig. VIP) shows that the 531 nm band is the most important spectral band for accurately identifying Chlorophyceae. In fact, at this wavelength, Chlorophyceae exhibits the highest reflectance among all other classes, highlighting the difference in accessories carotenoid proportion between seagrasses and green algae (@repolho2017seagrass).
